Generative AI for Image Reconstruction with Intensity Interferometry: A First Attempt

### 2024-12-09 Phase_II/draft/ParaGan.tex
modification

\section{The Parameter Analysis for GAN}
\begin{figure*}
	\centering
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_0.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_16.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_35.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_38.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_42.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_47.png}
	\end{subfigure}
	\caption{This set of figures shows the result of GAN for II. The left panels are signals using six baselines, which work as input for GAN. The first middle panel is the real image, also called ground truth, and GAN tries to mimic it during the training. The second middle panel is the reconstructed image, also called the predicted image, with trained GAN, and the end panel is the difference between the ground truth and the predicted image.}
	\label{fig:GAN}
\end{figure*}
Here, we discuss the parameters for the GAN structure, which is trained to reconstruct the image of stellar objects using II. Due to the adversarial nature of GAN, where the generator and discriminator compete in a min-max game, careful tuning of key parameters ensures that both networks converge properly.

\subsection{Data preparation}
First, we simulate fast-rotating stars, which are modeled as oblate spheroids with different radii and oblateness between 0.5 and 1. Different viewed angles have also been considered, with the effect of gravity darkening being linear dependence. The traced ellipses arise when integrating over the hour angle of the source. For the hyperparameter tuning and comparison of different telescopes, the total hour angle is about 11.5 hours. The ellipses are plotted and converted into grayscale images, resized, and stored as pure arrays to simplify further analysis. 
\begin{figure*}
	\centering
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom0.png}
		\caption{The monopole.}
		\label{fig:mom1}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom1.png}
		\caption{The centroids along x direction.}
		\label{fig:mom2}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom2.png}
		\caption{The centroids along y direction.}
		\label{fig:mom3}
	\end{subfigure}\hfill
	\caption{This set of figures shows the comparison of monopole, x-centroid, and y-centroid for ground truth and predicted images generated by trained GAN.}
	\label{fig:cen}
\end{figure*}
\begin{figure*}
	\centering
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom3.png}
		\caption{The 2nd order moment $\mu_{11}$.}
		\label{fig:mom4}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom4.png}
		\caption{The 2nd order moment $\mu_{20}$.}
		\label{fig:mom5}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom5.png}
		\caption{The 2nd order moment $\mu_{02}$.}
		\label{fig:mom6}
	\end{subfigure}\hfill
	\caption{The second-order central moments provide information about the size and shape of stellar objects. This set of figures shows all the second-order central moments for ground truth and predicted images generated by trained GAN.}
	\label{fig:struc}
\end{figure*}
\begin{figure*}
	\centering
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom6.png}
		\caption{The 3rd order moment $\mu_{30}$.}
		\label{fig:mom7}
	\end{subfigure}\hfill
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom7.png}
		\caption{The 3rd order moment $\mu_{03}$.}
		\label{fig:mom8}
	\end{subfigure}\hfill
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom8.png}
		\caption{The 3rd order moment $\mu_{21}$.}
		\label{fig:mom9}
	\end{subfigure}\hfill
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom9.png}
		\caption{The 3rd order moment $\mu_{12}$.}
		\label{fig:mom10}
	\end{subfigure}\hfill
	\caption{This set of figures shows all the 3rd-order central moments for ground truth and predicted images generated by trained GAN. It calculates the skewness and evaluates the brightness distribution of images.}
	\label{fig:moments}
\end{figure*}

First, Salt and Pepper noise is introduced at a rate of $\alpha$ (usually 0.5\%). Then, the images are resized, and the mean is subtracted. A two-dimensional Fast Fourier Transform, as well as a Fourier shift, is applied, which yields a complex number of every pixel. As II does not measure the phase, the absolute value is calculated (shown in Fig.~\ref{fig:ft} on a linear and logarithmic scale for visualization). The sparse sampling is also introduced, namely by a pixel-wise multiplication between the absolute valued Fourier transformed image and the sparse sampling map. The result is a map in the Fourier plane with several ellipses, also called the sparse sampling map. Fig.~\ref{fig:ft_base} shows the sparse sampling for Fig.~\ref{fig:image} as the source with four telescopes (Fig.~\ref{fig:teles}). Finally, the pixels are normalized and converted to 8-bit integers. This image represents the sparsely sampled complex visibility, as it can be measured with II. The image shown in Fig.~\ref{fig:ft_base} is the input image for the GAN, which also requires the ground truth image. So, the simulated stars are also resized using the same algorithm and converted to 8-bit integers to reduce the bias. The GAN must know the ground truth corresponding to an input image. Hence, they are merged side-by-side, shown in Fig.~\ref{fig:GANinput}, and used as input to train the GAN. This procedure is conducted for all the simulated stars, with 10\% acting as test data, 10\% as validation data, and the remaining 80\% as training data.

\subsection{GAN architecture}
The GAN used in this work is based on pix2pix \citep{isola2017image}, which uses a cGAN discussed in the last section. It is very robust and could already be applied to various problems. An example can be found in the TensorFlow tutorials\footnote{\url{https://www.tensorflow.org/tutorials/generative/pix2pix}}, where it is applied to a data set of architectural facades. However, to adapt the pix2pix GAN for the phase retrieval problem, some modifications are necessary. The network uses the TensorFlow \citep{abadi2016tensorflow} library, calculations are performed with scipy \citep{virtanen2020scipy}, and plots are drawn with matplotlib \citep{4160265}.
\begin{figure*}
	\includegraphics[width=\linewidth]{fig/reconstruction.png}
	\caption{The elliptical image of ground truth and predicted image by trained GAN. The elliptic curve (in red) was drawn using the second-order moment, and cyan dots show the centroids (mx, my) of the images. There are two arrows in each image showing the skewness of pixel distribution along the x and y axes (magenta and blue color).}
	\label{fig:recons}
\end{figure*}
\subsection{Hyperparameter tuning}
The GAN used for this work depends on several parameters, which will be explained here briefly (an in-depth discussion can be found in \citep{murphy2022probabilistic}).

The learning rate of the optimizer is a measure of how much the model changes per iteration. Small learning rates can lead to underfitting, and large learning rates can make the model unstable. It is therefore important to carefully choose the learning rate \citep{murphy2022probabilistic}. Fig.~\ref{fig:Plot_learning_rate_loss} shows the effect of the learning rate on both generator and discriminator loss for three different values of it. Both loss functions have fewer outliers with decreasing learning rates, which is expected, as there is less change in the models with low learning rates. Even though all models stabilize at the same level, low learning rates are preferred.

The kernel size is the size of the kernel in the convolutional layer. It determines the amount of pixels that are linearly combined into a new pixel. A large kernel size allows one to catch features that are a few pixels away, but it also means that unrelated features might be introduced. The kernel size does not have a big effect on the loss functions, as shown in Fig.~\ref{fig:Plot_kernel_size_loss}. Smaller kernel sizes seem to have more outliers, which means that either generator or discriminator gains an advantage. Therefore, a kernel size of 5 is preferred.

@@

The batch size defines the number of images that pass the network simultaneously. It is observed that smaller batch sizes can lead to better generalization \citep{prince2023understanding}. Fig.~\ref{fig:Plot_batchsize_loss} shows two different batch sizes. Here, when two images are processed simultaneously, there are fewer outliers, but since it increases the training time heavily, a batch size of 1 is used.

While training GAN, one can give the discriminator an advantage by increasing the number of training steps before returning to the generator training. It can increase the performance of the model. However, it increases the training time as well and leads to a lower discriminator loss score, as shown in Fig.~\ref{fig:Plot_discrep_loss}. At the same time, the generator score increases slightly, which should be no surprise. Because the generated images did not improve with higher discriminator training, they are usually trained in the same amount as the generator.

Finally, the amount of sparse sampling can be varied, giving the model access to more pixels. More telescopes automatically lead to more baselines and, hence, more pixels that are available. Fig.~\ref{fig:Plot_telescopes_loss} shows the loss functions for different numbers of telescopes. There is a huge disparity, which can partly be explained by the fact that the relation between telescopes and baselines is not linear. The Fourier plane can be sampled sixfold when using four telescopes in contrast to only two telescopes. In the latter case, both the generator and discriminator are not trained smoothly, as seen by the outliers. It is already better with three telescopes and very promising when using four telescopes. The degree of sparse sampling seems to have the biggest effect of all hyperparameters.

>>>>

\section{The Parameter for GAN Structure}

\begin{figure*}
	\centering
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_0.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_16.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_35.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_38.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_42.png}
	\end{subfigure}
	\begin{subfigure}{\linewidth}
		\includegraphics[width=\linewidth]{fig/testing_image/image_47.png}
	\end{subfigure}
	\caption{This set of figures shows the result of GAN for II. The left panels are signals using six baselines, which work as input for GAN. The first middle panel is the real image, also called ground truth, and GAN tries to mimic it during the training. The second middle panel is the reconstructed image, also called the predicted image, with trained GAN, and the end panel is the difference between the ground truth and the predicted image.}
	\label{fig:GAN}
\end{figure*}
Here, we discuss the parameters for the GAN structure, which is trained to reconstruct the image of stellar objects using II. Due to the adversarial nature of GAN, where the Generator and Discriminator compete in a min-max game, careful tuning of key parameters ensures that both networks are in good term to train the model.

\subsection{Data Preparation}
First, we simulate fast-rotating stars, which are modeled as oblate spheroids with different radii and oblateness between 0.5 and 1. Different viewed angles have also been considered, with the effect of gravity darkening being linear dependence. The traced ellipses arise when integrating over the hour angle of the source. For the hyperparameter tuning and comparison of different telescopes, the total hour angle is about 11.5 hours. The ellipses are plotted and converted into grayscale images, resized, and stored as pure arrays to simplify further analysis. 
\begin{figure*}
	\centering
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom0.png}
		\caption{The monopole.}
		\label{fig:mom1}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom1.png}
		\caption{The centroids along x direction.}
		\label{fig:mom2}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom2.png}
		\caption{The centroids along y direction.}
		\label{fig:mom3}
	\end{subfigure}\hfill
	\caption{This set of figures shows the comparison of monopole, x-centroid, and y-centroid for ground truth and predicted images generated by trained GAN.}
	\label{fig:cen}
\end{figure*}
\begin{figure*}
	\centering
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom3.png}
		\caption{The 2nd order moment $\mu_{11}$.}
		\label{fig:mom4}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom4.png}
		\caption{The 2nd order moment $\mu_{20}$.}
		\label{fig:mom5}
	\end{subfigure}\hfill
	\begin{subfigure}{0.33\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom5.png}
		\caption{The 2nd order moment $\mu_{02}$.}
		\label{fig:mom6}
	\end{subfigure}\hfill
	\caption{The second-order central moments provide information about the size and shape of stellar objects. This set of figures shows all the second-order central moments for ground truth and predicted images generated by trained GAN.}
	\label{fig:struc}
\end{figure*}
\begin{figure*}
	\centering
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom6.png}
		\caption{The 3rd order moment $\mu_{30}$.}
		\label{fig:mom7}
	\end{subfigure}\hfill
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom7.png}
		\caption{The 3rd order moment $\mu_{03}$.}
		\label{fig:mom8}
	\end{subfigure}\hfill
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom8.png}
		\caption{The 3rd order moment $\mu_{21}$.}
		\label{fig:mom9}
	\end{subfigure}\hfill
	\begin{subfigure}{0.50\linewidth}
		\includegraphics[width=\linewidth]{fig/moments/mom9.png}
		\caption{The 3rd order moment $\mu_{12}$.}
		\label{fig:mom10}
	\end{subfigure}\hfill
	\caption{This set of figures shows all the 3rd-order central moments for ground truth and predicted images generated by trained GAN. It calculates the skewness and evaluates the brightness distribution of images.}
	\label{fig:moments}
\end{figure*}

First, Salt and Pepper noise is introduced at a rate of $\alpha$ (usually 0.5\%). Then, the images are resized, and the mean is subtracted. A two-dimensional Fast Fourier Transform, as well as a Fourier shift, is applied, which yields a complex number of every pixel. As II does not measure the phase, the absolute value is calculated (shown in Fig.~\ref{fig:ft} on a linear and logarithmic scale for visualization). The sparse sampling is also introduced, namely by a pixel-wise multiplication between the absolute valued Fourier transformed image (Fig.~\ref{fig:ft}) and the sparse sampling map (Fig.~\ref{fig:base}). The result is a map in the Fourier plane with several ellipses, also called the sparse sampling map (Fig.~\ref{fig:ft_base}). It shows the sparse sampling for Fig.~\ref{fig:image} as the source with four telescopes (Fig.~\ref{fig:teles}). Finally, the pixels are normalized and converted to 8-bit integers. This image represents the sparsely sampled complex visibility, as it can be measured with II. The image shown in Fig.~\ref{fig:ft_base} is the input image for the GAN, which also requires the ground truth image. So, the simulated stars are also resized using the same algorithm and converted to 8-bit integers to reduce the bias. The GAN must know the ground truth corresponding to an input image. Hence, they are merged side-by-side, shown in Fig.~\ref{fig:GANinput}, and used as input to train the GAN. This procedure is conducted for all the simulated stars, with 10\% acting as test data, 10\% as validation data, and the remaining 80\% as training data.

\subsection{GAN Architecture}
The GAN used in this work is based on pix2pix, which uses a cGAN discussed in the last section \cite{isola2017image}. It is very robust and could already be applied to various problems. An example can be found in the TensorFlow tutorials\footnote{\url{https://www.tensorflow.org/tutorials/generative/pix2pix}}, where it is applied to a data set of architectural facades. However, to adapt the pix2pix GAN for the phase retrieval problem, some modifications are necessary. The network uses the TensorFlow library \cite{abadi2016tensorflow}, calculations are performed with scipy \cite{virtanen2020scipy}, and plots are drawn with matplotlib \cite{4160265}.

\subsection{Hyperparameter Tuning}
The GAN used for this work depends on several parameters, which will be explained here briefly (an in-depth discussion can be found in \cite{murphy2022probabilistic}).

The learning rate of the optimizer is a measure of how much the model changes per iteration. Small learning rates can lead to underfitting, and large learning rates can make the model unstable. It is therefore important to carefully choose the learning rate \cite{murphy2022probabilistic}. Fig.~\ref{fig:Plot_learning_rate_loss} shows the effect of the learning rate on both Generator and Discriminator loss for three different values of it. Both loss functions have fewer outliers with decreasing learning rates, which is expected, as there is less change in the models with low learning rates. Even though all models stabilize at the same level, low learning rates are preferred.

The kernel size is the size of the kernel in the convolutional layer. It determines the amount of pixels that are linearly combined into a new pixel. A large kernel size allows one to catch features that are a few pixels away, but it also means that unrelated features might be introduced. The kernel size does not have a big effect on the loss functions, as shown in Fig.~\ref{fig:Plot_kernel_size_loss}. Smaller kernel sizes seem to have more outliers, which means that either Generator or Discriminator gains an advantage. Therefore, a kernel size of 5 is preferred.

@@

The batch size defines the number of images that pass the network simultaneously. It is observed that smaller batch sizes can lead to better generalization \cite{prince2023understanding}. Fig.~\ref{fig:Plot_batchsize_loss} shows two different batch sizes. Here, when two images are processed simultaneously, there are fewer outliers, but since it increases the training time heavily, a batch size of 1 is used.

While training GAN, one can give the Discriminator an advantage by increasing the number of training steps before returning to the Generator training. It can increase the performance of the model. However, it increases the training time as well and leads to a lower Discriminator loss score, as shown in Fig.~\ref{fig:Plot_discrep_loss}. At the same time, the Generator score increases slightly, which should be no surprise. Because the generated images did not improve with higher Discriminator training, they are usually trained in the same amount as the Generator.

Finally, the amount of sparse sampling can be varied, giving the model access to more pixels. More telescopes automatically lead to more baselines and, hence, more pixels that are available. Fig.~\ref{fig:Plot_telescopes_loss} shows the loss functions for different numbers of telescopes. There is a huge disparity, which can partly be explained by the fact that the relation between telescopes and baselines is not linear. The Fourier plane can be sampled sixfold when using four telescopes in contrast to only two telescopes. In the latter case, both the Generator and Discriminator are not trained smoothly, as seen by the outliers. It is already better with three telescopes and very promising when using four telescopes. The degree of sparse sampling seems to have the biggest effect of all hyperparameters.

### 2024-12-09 Phase_II/draft/Result.tex
modification

\section{The Reconstructed Image with GAN}
In this section, we will first discuss phase retrieval using hyperparameters, as mentioned earlier, followed by an analysis where multiple sources are trained simultaneously. The best performance for image reconstruction has been observed with a learning rate of $2 \cdot 10^{-4}$, a kernel size of 5x5, and equal noise percentages in original and generated images. The batch size selected is 1, and the discriminator-generator receives equal training.
\subsection{The Predicted Image with Trained GAN}
The success of the GAN in training the model for Intensity Interferometry to reconstruct images of fast-rotating stars is demonstrated in Fig.~\ref{fig:GAN}. The GAN was trained on training datasets for 60,000 steps. After training, the GAN was tested on different validation datasets, producing predicted images of a fast-rotating star. Fig.~\ref{fig:GAN} presents a set of four combined images demonstrating the GAN's performance in reconstructing the shape, size, and brightness distribution of the fast-rotating star using II. The left panel shows the signals covered by six baselines, which serve as the input for the generator to train the GAN. The first, middle panel displays the real image, or ground truth, which the discriminator uses to differentiate between real and generated images from the generator. During training, the GAN aims to mimic these ground truth images. The second middle panel depicts the reconstructed image, or predicted image, produced by the trained GAN. This panel illustrates the success of the GAN in image reconstruction using II. The right panel shows the difference between the ground truth and the predicted image. This difference should be minimized for high-precision image reconstruction with the application of GAN on II. The predicted images of Fig.~\ref{fig:GAN} showed positive results. It accurately provides visual information about the source's size, shape, and brightness distribution over the surface. It is achieved here using only six baselines. However, the results can be further improved by increasing the number of telescopes to cover maximum (u, v) planes, making the existing and upcoming Cherenkov Telescope Array Observatory (CTAO) an ideal candidate. 
\subsection{Evaluation of GAN using moments}
The reconstructed images are good enough visually to claim the success of GAN in reconstructing the image using II. However, there is a need for statistical evaluation of this predicted image to guarantee success. So, we use image moments as a statistical method. Image moments are statistical properties that provide information about the reconstructed shape, size, and intensity distribution of the objects in the image. These moments help to quantify key features like the position, orientation, and distribution of brightness in the image, allowing for an objective evaluation of how well the predicted image corresponds to the actual object. We will analyze the consistency and accuracy of the GAN-generated image by comparing its moments to those of the ground truth. It provides a reliable framework for evaluating reconstruction quality, as image moments highlight the subtle differences in geometric and intensity properties that might not be visually apparent.

@@

The center of mass for the fast-rotating star or any other stellar object is calculated using the centroid (x-centroid and y-centroid). It represents the spatial position of the image and is calculated using first-order raw moment and monopole. The formulation of centroid along the x and y directions is
\begin{equation}
	\begin{aligned}
		&m_x = \frac{\sum_{x} x I(x,y)}{\sum_{x} \sum_{y} I(x, y)} = \frac{M_{10}}{M_{00}} \\
		&m_y = \frac{\sum_{y} y I(x,y)}{\sum_{x} \sum_{y} I(x, y)} = \frac{M_{01}}{M_{00}}
	\end{aligned}  
\end{equation}
Fig.~\ref{fig:mom2} and Fig.~\ref{fig:mom3} show the comparison of the x-centroid and y-centroid for 50 predicted images with respect to ground truths, respectively. The clustering of centroids in a given scale range for all the results explains that the reconstructed image correctly represents the spatial location of the fast-rotating star in compare with the ground truth.

@@

\subsection{The reconstructed Parameters for object}
The centroids $(m_x, m_y)$ represent the center of the fast-rotating star and spatial location on the image only. However, the calculated second-order central moment defines the orientation, semi-major axis, and eccentricity with respect to the center of the source \citep{teague1980image}. These parameters based on the moments fully describe the two-dimensional ellipse that fits the image data.

The orientation of a fast-rotating star along the line of sight is defined in terms of second-order central moments as
\begin{equation}
	\theta = \frac{1}{2}\arctan \big(\frac{2\mu_{11}}{\mu_{20} - \mu_{02}}\big).
	\label{eqn:orn}
\end{equation}
The semi-major and semi-minor axis of the stellar object will be calculated with the help of second-order central moment again and denoted as a and b here, respectively
\begin{equation}
	\begin{aligned}
		&a = 2\sqrt{mp + \delta} \\
		&b = 2\sqrt{mp - \delta}
	\end{aligned}
	\label{eqn:semi}
\end{equation}
where,
\begin{equation}
	mp = \frac{\mu_{20} + \mu_{02}}{2}
	\label{eqn:mp}
\end{equation}
and
\begin{equation}
	\delta = \frac{\sqrt{4\mu_{11}^2 + (\mu_{20} - \mu_{02})^2}}{2}.	
	\label{eqn:delta}
\end{equation}
With the help of the axis value, the eccentricity of the fast-rotating star is calculated as
\begin{equation}
	e = \sqrt{1 - a/b}.
	\label{eqn:eccen}
\end{equation}
Fig.~\ref{fig:recons} shows all zeroth to third order moments only for one ground and predicted image. In this figure, there is a red curve for both ground truth and predicted image, which is the representation of all second-order moments in terms of Eqn.~\ref{eqn:orn} to Eqn.~\ref{eqn:eccen} as orientation, axis value, and eccentricity. There is a cyan dot point as well, which represents the zeroth order moment for both images. The third-order moment is expressed in terms of arrows in magenta and blue color to view the skewness along the x and y axes.

>>>>

\section{The Reconstructed Image with GAN}

In this section, we will first discuss phase retrieval using hyperparameters, as mentioned earlier, followed by an analysis where multiple sources are trained simultaneously. The best performance for image reconstruction has been observed with a learning rate of $2 \cdot 10^{-4}$, a kernel size of 5x5, and equal noise percentages in original and generated images. The batch size selected is 1, and the Discriminator-Generator receives equal training.

\subsection{The Predicted Image with Trained GAN}
The success of the GAN in training the model for Intensity Interferometry to reconstruct images of fast-rotating stars is demonstrated in Fig.~\ref{fig:GAN}. The GAN was trained on training datasets for 60,000 steps. After training, the GAN was tested on different validation datasets, producing predicted images of a fast-rotating star. Fig.~\ref{fig:GAN} presents a set of four combined images demonstrating the GAN's performance in reconstructing the shape, size, and brightness distribution of the fast-rotating star using II. The left panel shows the signals covered by six baselines, which serve as the input for the Generator to train the GAN. The first middle panel displays the real image, or ground truth, which the Discriminator uses to differentiate between real and generated images from the Generator. During training, the GAN aims to mimic these ground truth images. The second middle panel depicts the reconstructed image, or predicted image, produced by the trained GAN. This panel illustrates the success of the GAN in image reconstruction using II. The right panel shows the difference between the ground truth and the predicted image. This difference should be minimized for high-precision image reconstruction with the application of GAN on II. The predicted images of Fig.~\ref{fig:GAN} showed positive results. It accurately provides visual information about the source's size, shape, and brightness distribution over the surface. It is achieved here using only six baselines. However, the results can be further improved by increasing the number of telescopes to cover maximum (u, v) planes, making the existing and upcoming Cherenkov Telescope Array Observatory (CTAO) an ideal candidate.

\subsection{Evaluation of GAN using Moments}
The reconstructed images are good enough visually to claim the success of GAN in reconstructing the image using II. However, there is a need for statistical evaluation of this predicted image to guarantee success. So, we use image moments as a statistical method. Image moments are statistical properties that provide information about the reconstructed shape, size, and intensity distribution of the objects in the image. These moments help to quantify key features like the position, orientation, and distribution of brightness in the image, allowing for an objective evaluation of how well the predicted image corresponds to the actual object. We will analyze the consistency and accuracy of the GAN-generated image by comparing its moments to those of the ground truth. It provides a reliable framework for evaluating reconstruction quality, as image moments highlight the subtle differences in geometric and intensity properties that might not be visually apparent.

@@

The center of mass for the fast-rotating star or any other stellar object is calculated using the centroid (x-centroid and y-centroid). It represents the spatial position of the image and is calculated using first-order raw moment and monopole. The formulation of centroid along the x and y directions is
\begin{equation}
	\begin{aligned}
		&m_x = \frac{\sum_{x} x I(x,y)}{\sum_{x} \sum_{y} I(x, y)} = \frac{M_{10}}{M_{00}} \\
		&m_y = \frac{\sum_{y} y I(x,y)}{\sum_{x} \sum_{y} I(x, y)} = \frac{M_{01}}{M_{00}}
	\end{aligned}  
\end{equation}
Fig.~\ref{fig:mom2} and Fig.~\ref{fig:mom3} show the comparison of the x-centroid and y-centroid for 50 predicted images with respect to ground truths, respectively. The clustering of centroids in a given scale range for all the results explains that the reconstructed image correctly represents the spatial location of the fast-rotating star compared with the ground truth.

@@

\subsection{The Reconstructed Parameters for Object}
The centroids $(m_x, m_y)$ represent the center of the fast-rotating star and spatial location on the image only. However, the calculated second-order central moment defines the orientation, semi-major axis, and eccentricity with respect to the center of the source \citep{teague1980image}. These parameters based on the moments fully describe the two-dimensional ellipse that fits the image data.

The orientation of a fast-rotating star along the line of sight is defined in terms of second-order central moments as
\begin{equation}
	\theta = \frac{1}{2}\arctan \big(\frac{2\mu_{11}}{\mu_{20} - \mu_{02}}\big).
	\label{eqn:orn}
\end{equation}
The semi-major and semi-minor axis of the stellar object will be calculated with the help of second-order central moment again and denoted as a and b here, respectively
\begin{equation}
	\begin{aligned}
		&a = 2\sqrt{mp + \delta} \\
		&b = 2\sqrt{mp - \delta}
	\end{aligned}
	\label{eqn:semi}
\end{equation}
where,
\begin{equation}
	mp = \frac{\mu_{20} + \mu_{02}}{2}
	\label{eqn:mp}
\end{equation}
and
\begin{equation}
	\delta = \frac{\sqrt{4\mu_{11}^2 + (\mu_{20} - \mu_{02})^2}}{2}.	
	\label{eqn:delta}
\end{equation}
With the help of the axis value, the eccentricity of the fast-rotating star is calculated as
\begin{equation}
	e = \sqrt{1 - a/b}.
	\label{eqn:eccen}
\end{equation}
Eqns.~\ref{eqn:orn}-\ref{eqn:eccen} depending on values explains the elliptical nature of the stellar object (here, fast-rotating star) and provides information on shape and size. However, the information about the brightness distribution is gathered using skewness, which can be understood using third and higher-order moments.

### 2024-12-09 Phase_II/draft/conclusion.tex
modification

\section{Conclusion}
Intensity Interferometry (II), a technique, is emerging after a long gap to overcome the challenges of Amplitude Interferometry with optical wavelength range. It is rejuvenated with support from advanced facilities at the Imaging Cherenkov Telescope Arrays (ICTAs). These telescopes are instrumental in capturing high-resolution images equipped with large aperture areas and sensitive photon detectors (capable of resolving signals within a nanosecond) \citep{dravins2013optical}. The Major Atmospheric Gamma Imaging Cherenkov Telescope (MAGIC), one of the CTA in the world, is now also being used for II, observing stellar objects with its 17-meter diameter mirror \citep{lorenz2004status}. The telescope has already demonstrated the potential of II by measuring and comparing the diameters of individual stars \citep{abe2024performance}. Other existing arrays (Very Energetic Radiation Imaging Telescope Array System (VERITAS) and High Energy Stereoscopic System (HESS)) are also making progress as II when gamma-ray observation stops \citep{kieda2021veritas, zmija2022optical}. Future observations, made possible by more advanced, sensitive instrumentation and upcoming CTAO, promise to push the boundaries of what we can observe with optical range. However, there is a drawback of II in terms of phase loss of signal, capturing only its magnitude via photon correlation. Here, the challenge of phase retrieval in Intensity Interferometry (II) has been effectively addressed through the application of machine learning techniques, particularly with conditional Generative Adversarial Networks (cGAN). Our study demonstrates that the application of cGAN on II data successfully recovers the size, shape, and brightness distribution of a fast-rotating star. The evaluation based on image moments, specifically the monopole, second, and third-order moments, supports the effectiveness of the cGAN in achieving accurate image reconstruction from a one-night simulation of II using six baselines. A critical factor in the performance of the reconstruction process is the extent of Fourier plane coverage, which is determined by the number of available telescopes and the total observing time. The brightness distribution can likely be reconstructed with even higher precision with a full night of observation using four telescopes. Future work could explore different observatory layouts to study the impact on image reconstruction quality, such as the Southern Cherenkov Telescope Array (CTA) with MAGIC.

>>>>

\section{Conclusion}
Intensity Interferometry (II), a technique, is emerging after a long gap to overcome the challenges of Amplitude Interferometry with optical wavelength range. It is rejuvenated with support from advanced facilities at the Imaging Cherenkov Telescope Arrays (ICTAs). These telescopes are instrumental in capturing high-resolution images equipped with large aperture areas and sensitive photon detectors (capable of resolving signals within a nanosecond) \cite{dravins2013optical}. The Major Atmospheric Gamma Imaging Cherenkov Telescope (MAGIC), one of the CTA in the world, is now being used for II, observing stellar objects with its 17-meter diameter mirror \cite{lorenz2004status}. The telescope has already demonstrated the potential of II by measuring and comparing the diameters of individual stars \cite{abe2024performance}. Other existing arrays (Very Energetic Radiation Imaging Telescope Array System (VERITAS) and High Energy Stereoscopic System (HESS)) are also making progress as II when gamma-ray observation stops \cite{kieda2021veritas, zmija2022optical}. Future observations, made possible by more advanced, sensitive instrumentation and upcoming CTAO, promise to push the boundaries of what we can observe with optical range. However, there is a drawback of II in terms of phase loss of signal, capturing only its magnitude via photon correlation. Here, the challenge of phase retrieval in II has been effectively addressed through the application of machine learning techniques, particularly with conditional Generative Adversarial Networks (cGAN). Our study demonstrates that the application of cGAN on II data successfully recovers the size, shape, and brightness distribution of a fast-rotating star. The evaluation based on image moments, specifically the monopole, second, and third-order moments, supports the effectiveness of the cGAN in achieving accurate image reconstruction from a one-night simulation of II using six baselines. A critical factor in the performance of the reconstruction process is the extent of Fourier plane coverage, which is determined by the number of available telescopes and the total observing time. The brightness distribution can likely be reconstructed with even higher precision with a full night of observation using four telescopes. Future work could explore different observatory layouts to study the impact on image reconstruction quality, such as the Southern Cherenkov Telescope Array (CTA) with MAGIC.
